---
title: "Congenial Imputation Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Congenial Imputation Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`ikunfrailty` targets the special analysis model derived in
`frailtypack_congenial_pdmi_only_technical_note.tex`. The package does not ask
users to write the disease part of the imputation model. Instead, it derives
the congenial kernel from the analysis formula and combines it with the
working covariate prior supplied by `pdmi_prior()`.

The imputation target is

\[
f(X_M \mid X_O,Y,P,A=1;\psi) \propto
\prod_i
\frac{s_i(Y_i,X_i)M_i(Y_i\mid X_i;\theta)}
{\alpha_i^F(X_i;\theta)}
g_i^A(X_i\mid P_i;\eta_X).
\]

## What the Formula Provides

The formula identifies:

- the survival response through `Surv()`;
- the event/status column;
- the two disease covariates;
- the family cluster through `cluster(family)`.

For example:

```{r formula, eval = FALSE}
Surv(t0, time, status) ~ mgene + newx + cluster(famID)
```

The first package engine validates that this is the V2.2 scalar target model:
one binary carrier covariate and one continuous PRS covariate. The completed
data analysis uses the frailtypack model

```{r frailtypack, eval = FALSE}
frailtypack::frailtyPenal(
  Surv(t0, time, status) ~ mgene + newx + cluster(famID),
  data = completed_data,
  hazard = "Weibull",
  RandDist = "LogN",
  covMatrix1 = K,
  recurrentAG = TRUE,
  proband = completed_data$proband,
  currentage = completed_data$currentage,
  agemin = 0
)
```

## What the Prior Provides

The user supplies only \(g_i^A(X_i \mid P_i; \eta_X)\):

```{r prior, eval = FALSE}
prior <- pdmi_prior(
  continuous = list(newx = normal_kinship(~ mgene, covariance = "kinship+iid")),
  binary = list(mgene = carrier_hwe(q = "estimate"))
)
```

This chooses the realistic prior versions `C-R`, `B-R`, or `J-R`, depending on
which variables are listed in `impute`. Known oracle versions can be requested
with a known kinship PRS prior and numeric carrier allele frequency.

## Derived Components

`pdmi_frailty()` derives:

- \(M_i(Y_i \mid X_i;\theta)\): the Weibull/lognormal correlated-frailty
  likelihood implied by the analysis formula;
- \(\alpha_i^F(X_i;\theta)\): the pop+ proband ascertainment probability using
  proband `currentage`, covariates, kinship diagonal, and quadrature;
- \(s_i(Y_i,X_i)\): the support restriction requiring one affected carrier
  proband per family;
- disease-parameter posterior draws by exact slice sampling on the reported
  omega scale;
- frailty draws by family-level elliptical slice sampling;
- continuous PRS missing-value draws from the conditional normal prior times
  the disease and ascertainment terms;
- binary carrier missing-value draws by Mendelian genotype Gibbs updates.

For joint missingness, the binary genotype update includes the continuous
prior log density when the PRS prior depends on carrier status.

Inspect the derived object:

```{r inspect, eval = FALSE}
fit <- pdmi_frailty(
  Surv(t0, time, status) ~ mgene + newx + cluster(famID),
  data = dat,
  kinship = K,
  impute = c("newx", "mgene"),
  prior = prior
)

imputation_model(fit)
```
