---
title: "ikunfrailty Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ikunfrailty Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This workflow shows the intended user path for the validated V2.2 scalar
engine: one continuous PRS covariate and one binary carrier covariate in an
ascertainment-corrected frailtypack correlated frailty model.

For a full example using the shipped V2.2-style simfam replicate with 498
families and 20% MAR missingness, see the installed example script
`system.file("examples", "v22-one-replicate-mi.R", package = "ikunfrailty")`
and the vignette "V2.2 One-Replicate Multiple-Imputation Example".

## Inputs

The analysis data must contain survival time/status columns, a family cluster,
an individual id, a proband indicator, current age, one continuous PRS
covariate, and one binary carrier covariate.

```{r setup, eval = FALSE}
library(ikunfrailty)
library(survival)

data("ikun_example_joint_mar20", package = "ikunfrailty")
data("ikun_example_kinship", package = "ikunfrailty")
data("ikun_example_pedigree", package = "ikunfrailty")

prior <- pdmi_prior(
  continuous = list(
    newx = normal_kinship(~ mgene, covariance = "kinship+iid")
  ),
  binary = list(
    mgene = carrier_hwe(q = "estimate", q0 = 0.02, n0 = 50)
  )
)
```

## Fit

`pdmi_frailty()` derives the disease component, ascertainment denominator,
support restriction, and sampler from the analysis formula. The user supplies
only the working covariate prior.

```{r fit, eval = FALSE}
fit <- pdmi_frailty(
  formula = Surv(t0, time, status) ~ mgene + newx + cluster(famID),
  data = ikun_example_joint_mar20,
  kinship = ikun_example_kinship,
  impute = list(continuous = "newx", binary = "mgene"),
  prior = prior,
  M = 20,
  B = 50,
  numit = 10,
  report = "both",
  pedigree = ikun_example_pedigree,
  seed = 20260527,
  progress = TRUE
)
```

Use `impute = "newx"` for continuous-only missingness, `impute = "mgene"` for
binary-only missingness, and both names for joint missingness. Binary and joint
missingness require observed proband carrier status under the pop+ support.

## Parameter Estimates

```{r parameters, eval = FALSE}
summary(fit)
coef(fit)
vcov(fit)
```

The reported parameter vector is
`log.rho`, `log.lambda`, `beta_b`, `beta_c`, and `sigma_u2`.

## Penetrance

By default, penetrance is evaluated at ages 40, 50, 60, 70, and 80; PRS
profiles -0.5, 0, and 0.5; carrier profiles 0 and 1; and `k0 = 1`.

```{r penetrance, eval = FALSE}
pen_summary(fit)
p <- pen_plot(fit)
p
```

You can override the grid:

```{r pen-grid, eval = FALSE}
fit2 <- pdmi_frailty(
  Surv(t0, time, status) ~ mgene + newx + cluster(famID),
  data = dat,
  kinship = K,
  impute = "newx",
  prior = prior,
  pen_grid = list(age = seq(35, 85, by = 5), prs = c(-1, 0, 1), gene = c(0, 1)),
  M = 20,
  B = 50
)
```

## Inspect the Derived Model

```{r kernel, eval = FALSE}
imputation_model(fit)
```

This prints the kernel components derived from the analysis formula plus
`prior`.
