---
title: "Priors and Arguments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Priors and Arguments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette summarizes the helper functions and the main arguments of
`pdmi_frailty()`.

## Prior Constructors

### Estimated continuous PRS prior

```{r cr, eval = FALSE}
normal_kinship(~ mgene, covariance = "kinship+iid")
```

This maps to the V2.2 `C-R` or `J-R` continuous prior. The mean is a linear
function of carrier status and the covariance is
\(\tau_g^2 K_i + \tau_e^2 I_i\), with nuisance parameters updated inside the
PDMI sampler.

### Known continuous PRS prior

```{r co, eval = FALSE}
normal_kinship(~ 1, covariance = "kinship", sigma2 = 0.1, estimate = FALSE)
```

This maps to `C-O` or `J-O`. The first validated engine uses the V2.2
mean-zero oracle PRS prior with covariance `sigma2 * K_i`.

### Estimated carrier prior

```{r br, eval = FALSE}
carrier_hwe(q = "estimate", q0 = 0.02, n0 = 50)
```

This maps to `B-R` or `J-R`. The allele frequency is updated under the
carrier-conditioned pop+ Mendelian/HWE prior.

### Known carrier prior

```{r bo, eval = FALSE}
carrier_hwe(q = 0.02)
```

This maps to `B-O` or `J-O`.

## Main Arguments

`formula` must contain a `Surv()` response, two scalar disease covariates, and
one `cluster()` term.

`data` is the analysis data. Required columns are:

- survival start/time/status columns;
- family cluster;
- individual id;
- proband indicator;
- current age;
- the continuous and binary covariates.

`kinship` is the analysis-record kinship matrix. Named matrices are matched to
`id`; unnamed matrices must already be in data order.

`impute` selects which covariates are subject to missingness:

```{r impute, eval = FALSE}
impute = "newx"
impute = "mgene"
impute = c("newx", "mgene")
impute = list(continuous = "newx", binary = "mgene")
```

`prior` must include prior components for the variables listed in `impute`.
For joint missingness, the first validated engine supports matched oracle
priors (`C-O + B-O`) or matched estimated priors (`C-R + B-R`).

`M` is the number of imputations.

`B` is the burn-in count. The V2.2 sampler runs `B + numit` updates and keeps
the final state for each imputation.

`pen_grid` can override the default penetrance grid:

```{r pen-grid, eval = FALSE}
pen_grid = list(
  age = c(40, 50, 60, 70, 80),
  prs = c(-0.5, 0, 0.5),
  gene = c(0, 1),
  k0 = 1
)
```

`control = pdmi_control()` changes numerical controls such as quadrature order,
frailtypack maximum iterations, and slice sampler settings.

## Unsupported First-Version Cases

The API is designed for later extensions, but version 0.0.1 errors clearly for
unsupported cases:

- more than one continuous covariate;
- more than one binary covariate;
- transformed or interaction disease covariate terms;
- Bernoulli GLM binary priors;
- multivariate continuous priors;
- missing proband carrier status.
