---
title: "V2.2 One-Replicate Multiple-Imputation Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{V2.2 One-Replicate Multiple-Imputation Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This vignette shows a hands-on workflow using the V2.2-style example data
shipped with `ikunfrailty`. The data were generated once from the original
V2.2 simfam/FamEvent pipeline and saved in the package so users do not need the
simulation source tree to try `pdmi_frailty()`.

The shipped replicate contains 498 selected pop+ families. Three incomplete
analysis data sets were made from the same complete replicate:

- `ikun_example_continuous_mar20`: 20% MAR missingness in `newx`;
- `ikun_example_binary_mar20`: 20% MAR missingness in non-proband `mgene`;
- `ikun_example_joint_mar20`: 20% MAR missingness in both `newx` and non-proband `mgene`.

The complete runnable script is installed with the package:

```{r script-location, purl = FALSE}
system.file("examples", "v22-one-replicate-mi.R", package = "ikunfrailty")
```

## Load Data

```{r load-data, purl = FALSE}
library(ikunfrailty)
library(survival)

data("ikun_example_complete", package = "ikunfrailty")
data("ikun_example_continuous_mar20", package = "ikunfrailty")
data("ikun_example_binary_mar20", package = "ikunfrailty")
data("ikun_example_joint_mar20", package = "ikunfrailty")
data("ikun_example_kinship", package = "ikunfrailty")
data("ikun_example_pedigree", package = "ikunfrailty")
data("ikun_example_missing_diagnostics", package = "ikunfrailty")
data("ikun_example_metadata", package = "ikunfrailty")

dim(ikun_example_complete)
length(unique(ikun_example_complete$famID))
ikun_example_missing_diagnostics
ikun_example_metadata
```

The package stores the aligned kinship matrix separately because
`pdmi_frailty()` needs it as the `kinship` argument. The pedigree copy is needed
for binary-only and joint missingness.

## Specify The Working Prior

Users provide only the working covariate prior. The disease likelihood,
ascertainment denominator, support restriction, and posterior-draw sampler are
derived from the analysis formula.

```{r prior, purl = FALSE}
prior <- pdmi_prior(
  continuous = list(
    newx = normal_kinship(~ mgene, covariance = "kinship+iid")
  ),
  binary = list(
    mgene = carrier_hwe(q = "estimate", q0 = 0.02, n0 = 50)
  )
)

analysis_formula <- Surv(t0, time, status) ~ mgene + newx + cluster(famID)
```

## Continuous Missingness

```{r fit-continuous, purl = FALSE}
fit_continuous <- pdmi_frailty(
  formula = analysis_formula,
  data = ikun_example_continuous_mar20,
  kinship = ikun_example_kinship,
  impute = list(continuous = "newx"),
  prior = prior,
  M = 20,
  B = 50,
  numit = 10,
  report = "both",
  pedigree = ikun_example_pedigree,
  seed = 910135,
  progress = TRUE
)

summary(fit_continuous)
```

## Binary Missingness

```{r fit-binary, purl = FALSE}
fit_binary <- pdmi_frailty(
  formula = analysis_formula,
  data = ikun_example_binary_mar20,
  kinship = ikun_example_kinship,
  impute = list(binary = "mgene"),
  prior = prior,
  M = 20,
  B = 50,
  numit = 10,
  report = "both",
  pedigree = ikun_example_pedigree,
  seed = 920135,
  progress = TRUE
)

summary(fit_binary)
```

## Joint Missingness

```{r fit-joint, purl = FALSE}
fit_joint <- pdmi_frailty(
  formula = analysis_formula,
  data = ikun_example_joint_mar20,
  kinship = ikun_example_kinship,
  impute = list(continuous = "newx", binary = "mgene"),
  prior = prior,
  M = 20,
  B = 50,
  numit = 10,
  report = "both",
  pedigree = ikun_example_pedigree,
  seed = 930135,
  progress = TRUE
)

summary(fit_joint)
pen_summary(fit_joint)
imputation_model(fit_joint)
pen_plot(fit_joint)
```

For a quick syntax smoke test, reduce `M`, `B`, and `numit`. The values above
match the V2.2 package-performance setting and therefore can take time on a
local laptop.
